---
title: "R Notebook final version"
output: html_notebook
---
# EDA of the Olist datasets
```{r library}
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape)
library(stringr)
library(heatmaply)

```

#1-Marketing Channel Effectiveness

Sellers on the Olist platform come from various marketing channels (social media, referals, organic searches etc…).We will have a look at which one was the most effective at generating leads and interest from potential sellers. Marketing Qualified Lead(MQL)' are potential reseller/manufacturer who have an interest in selling their products on Olist.

```{r datasets import }
mql = read.csv("../data/olist_marketing_qualified_leads_dataset.csv")
#parse first contact as Date
mql$first_contact_date = as.Date(mql$first_contact_date,tz="UTC" ,format="%Y-%m-%d")

#create new date colmn in ormat y-m
mql$date_ym = format(as.Date(mql$first_contact_date), format="%Y-%m")

#replace empty cells with NA in each column
mql_upd = mql %>% 
  mutate(origin = na_if(origin, ""))
#check the number of NA
colSums(is.na(mql_upd))
```

there are 60 NA values in origin, we will replace those values with "others"
```{r}
#group all blank cells into "others"
mql_2 = mql_upd %>% mutate(origin = replace_na(origin, "others"))

#now remove mql_upd
#rm(mql_upd)

#create a new dataframe with montly count of mql_id 
mql_monthly = mql_2 %>% 
  group_by(date_ym) %>%
  summarise(mql_id_count = n())

#plot the volume of mql over the dates
ggplot(data=mql_monthly,aes(x=date_ym,y=mql_id_count,group=1)) +
  geom_line(color="blue")+
  geom_point()+
  labs(title="MQL Volume (Jun 2017 - May 2018)", fontsize=14)

```
Something happened in January 2018, which made the volume of MQL soar to over
1000, it has been on an upward trajectory since then

#MQL Volume by Marketing Channel
* The type of marketing channel is recorded in 'origin' field.
```{r}

#new dataframe
mql_origin = subset(mql_2, select=c(date_ym,origin))

#create a new column for count
mql_origin = mql_origin %>% 
  group_by(date_ym,origin) %>%
  summarise(mql_id_count = n())

#visualize with heat map
ggplot(data = mql_origin,aes(x=date_ym,y=origin)) +
  geom_tile(aes(fill=mql_id_count),show.legend = TRUE)+
  scale_fill_distiller(palette = "YlGnBu",direction=1)+
  theme_light() +
  theme(axis.text.x = element_text(angle=90,vjust=0.5,hjust=1)) #rotate x-axis t

```
-The channels that generate the most lead are organic searches followed by 
paid-searches.
- Since April 2018, the channel 'social' has generated more or less the same as 
paid searches.
-Should the cost of paid search increase, it would be wise to look at social 
media as an alternative.

#2. Sales Performance Overview
* A MQL who finally signed up for seller is called a closed deal.

Once a MQL has filled a form on one of the landing pages to sign up to be a 
seller on the platform, a Sales Development Representative(SDR) get in touch with
the MQL and gathers more information about the lead. 
A Sale representative(SR) then discuss with the lead. The interaction between 
SDRs/SRs and MQLs is one of the things that determine whether a MLQ sign up to 
become a seller on the platform.

In the following section I will explore two aspects of sales result: conversion rate and sales length.
```{r}
cl_deals = read.csv("../data/olist_closed_deals_dataset.csv")

#parsing won_date as datetime
cl_deals$won_date = as_datetime(cl_deals$won_date, format="%Y-%m-%d %H:%M:%S")

#merging this dataset and the MQL dataset
merged_data = merge(mql,cl_deals,by="mql_id",all.x = TRUE) #all.x=TRUE is left.Join()

#replace all empty cells in cl_deals and merged_data by NA

cl_deals_upd = subset(cl_deals,select=-c(won_date)) %>% na_if("")
cl_deals_upd$won_date = cl_deals$won_date

merged_data_upd = merged_data[,c(1:8,10:18)] %>% na_if("")
merged_data_upd$won_date = merged_data$won_date

#In merged_data replace all NA in origin with "others"
merged_data_upd = merged_data_upd %>% mutate(origin = replace_na(origin, "others"))

 
```

```{r}
#create a new column that holds true if the MLQ signs in,  and false otherwise
merged_data_upd$sign_in_bool = !is.na(merged_data_upd$seller_id)

#compute count of monthly closed_deals 
monthly_cl = merged_data_upd %>% 
  subset(select=c(date_ym,sign_in_bool)) %>% 
  group_by(date_ym) %>% 
  summarise(monthly_dl_closed = sum(sign_in_bool))

#plot the relationship

ggplot(data=monthly_cl, aes(x=date_ym,y=monthly_dl_closed,group=1))+
  geom_line(color="blue")+
  geom_point()+
  labs(title='Closed Deal Volume (Jun 2017 - May 2018)',y="signing", fontsize=14)+
  theme_light() +
  theme(axis.text.x = element_text(angle=90,vjust=0.5,hjust=1))
```
similar to the MQL lead volume, the volume of sign ins has drastically increased 
starting january 2018

#How many of the MQL ended up signing to become sellers?
```{r}
signing_rate = merged_data_upd %>% 
  subset(select=c(date_ym,sign_in_bool)) %>% 
  group_by(date_ym) %>% 
  summarise(monthly_dl_closed = sum(sign_in_bool),monthly_dl_count=n()) %>% 
  mutate(signing_rate_in_percentage=100*(monthly_dl_closed/monthly_dl_count))

#round the percentage values to 1 decimal place
signing_rate$signing_rate_in_percentage =
  round(signing_rate$signing_rate_in_percentage, digits=1)

#visualizing the relationship
ggplot(data=signing_rate, aes(x=date_ym,y=signing_rate_in_percentage,group=1))+
  geom_line(color="blue")+
  geom_point()+
  labs(title='Conversion Rate (Jun 2017 - May 2018)',y="Signing rate in %", fontsize=14)+
  theme_light() +
  theme(axis.text.x = element_text(angle=90,vjust=0.5,hjust=1))
```
#Digging into Closed Deal¶
Before move on to the next subject, I will explore closed deals in more depth to see whether there is room for improvement in sales process. Specifically looking at how behaviour profile affect the chance of closing a deal

#Attributes of a closed deals
```{r}
#a look at the differents type of behavior profile of the MQL
cd_profile  = subset(cl_deals_upd, is.na(cl_deals_upd$lead_behaviour_profile) ==FALSE)

cd_profile %>% 
  group_by(lead_behaviour_profile) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

# Combine the four types of mixed profiles(2.4%) into 'others'
mixed_profiles =list('cat, wolf', 'eagle, cat', 'eagle, wolf', 'shark, cat', 'shark, wolf')

  cd_profile$lead_behaviour_profile[cd_profile$lead_behaviour_profile %in% mixed_profiles] <- 'other profiles'
  
#look at the profiles and the count of seller id associated with each
  cd_profile %>% 
    group_by(lead_behaviour_profile) %>% 
    summarise(count = n())
#plot this
cd_profile %>% 
    group_by(lead_behaviour_profile) %>% 
    summarise(count = n()) %>% 
  ggplot(aes(y= forcats:: fct_reorder(lead_behaviour_profile,count) , x=count)) +
  geom_col(fill="purple") +
  labs(y="behaviour profile of the lead")
```
#Sales reps performance analysis

```{r}

cd_profile_tb = cd_profile %>% group_by(sr_id,lead_behaviour_profile) %>% 
  summarise(number_of_sales = n())

cd_profile_wd = pivot_wider(cd_profile_sr,names_from = lead_behaviour_profile , values_from = number_of_sales,values_fill = 0)

#visualizing the table
ggplot(data = cd_profile_tb,aes(y=sr_id,x=lead_behaviour_profile)) +
  geom_tile(aes(fill=number_of_sales),show.legend = TRUE)+
  theme(panel.background = element_rect(color="blue"))


```
SR at row 8 was the highest performer during the period and should be matched 
with cat, eagle and wolf MQL

Sales rep at rows 12,21 should be matched with shark MQL

sales rep at rows 8,12,10,20 have shown good results with wolf MQL, and should be matched with those MQL who fall under the wolf profile.

Clearly most of our sales were done with cat MQL, however sales rep at rows 13,18,11,3,15,5 have consistently performed rather poorly even with the "easy" cat 
MQL, This should be looked into to determine the possible reasons fr such poor performance

On the other hand sales rep at rows 8,12,21,4,10 have performed very well.





